---
title: "biostat776-project1-Shengyuan-Luo"
author: "davidluoshy"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project 1 - setup

Here I load packages "here" and "tidyverse"

```{r}
library("here")
library("tidyverse")
```

Next I create a directory named "data." The if statement helps avoid re-creating the directory if it already exists.

```{r}
if (!dir.exists(here("data"))) {
    dir.create(here("data"))
}
```

I use the following code to download the dataset. Again, the if statement checks if the dataset already exists. If so, data will not be re-downloaded.

```{r}
if (!file.exists(here("data", "chocolate.RDS"))) {
    url_csv <- "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-01-18/chocolate.csv"
    chocolate <- readr::read_csv(url_csv)

    # save the file to RDS objects
    saveRDS(chocolate, file = here("data", "chocolate.RDS"))
}
```

Making the dataset a tibble and taking a glimpse at the data.

```{r}
chocolate <- readRDS(here("data", "chocolate.RDS"))
as_tibble(chocolate)
glimpse(chocolate)
```

## Part 1: 

## Part 2:

## Part 3:

## Part 4:

## Part 5:

## Part6: Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
